---
title: "veterinary-perfusion-analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{veterinary-perfusion-analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Setup

First, let us load all the packages we will need.

```{r setup}
library(vetperf)
library(ggplot2)
library(data.table)
library(lme4)
```

# Extract Perfusion parameters

Quickly extract perfusion parameters from measurements.

## Data Import

First let us load the data. The vetperf package provides a dataset with measurements from 20 dogs, but you
can also easily load your own by calling the "fread" function from the "data.table" package.

```{r}
# Alternatively load your own data
# animals = fread('path/to/your/data.csv', stringsAsFactors = TRUE)
head(animals, 10)
```

## Compute parameters

Call the main function to compute perfusion parameters. Here, we let internally use cross-validation
to find the best hyperparameters and return normalized perfusion parameters. We use the "cs" (corresponding for the Centrum Semiovale) ROI for the normalization. In addition, we remove all measurements for which
no perfusion parameters could be computed (i.e. if the algorithm does not converge).

```{r}
perfdata = vetperf.data(animals, arrival_frame = 9, delta_t = 1.6, echo_time = 0.03, baseroi = "cs")
perfdata = na.omit(perfdata)
```

# Analyze disease influence

Let us briefly describe how we can analyze the perfusion parameters. First, we need to add the information
which dogs were healthy or not (not included in the dataset for simplicity).

## Format data

The first 8 dogs were healthy, the last 12 were diseased.

```{r}
perfdata[animal %in% c(1:8), state := "healty"]
perfdata[animal %in% c(9:20), state := "diseased"]
perfdata$state = as.factor(perfdata$state)
```

## Plot densities

Quickly plot the densities of parameters to visualy identify differences between healthy and diseased dogs.
The most obvious difference is seen for the Hippocampus.

```{r}
ggplot(perfdata) + aes(x = MTT, color = state) + geom_density() + facet_wrap(~roi)
ggplot(perfdata) + aes(x = CBF, color = state) + geom_density() + xlim(0,5.) + facet_wrap(~roi)
ggplot(perfdata) + aes(x = CBV, color = state) + geom_density() + xlim(0,5.) + facet_wrap(~roi)
```

## Check for significance

Fit Linear Mixed Models to assess the significance of the factor "state". We start with the MTT parameter.
First, fit a model not including the factor and then one including the factor. Finally, we compare
both models using a $\chi_2$ test. The result clearly indicates that the factor "state" is significant.

```{r}
model.mtt.0 = lmer(MTT ~ roi + (1|animal), data=perfdata)
model.mtt.1 = lmer(MTT ~ roi*state + (1|animal), data=perfdata)
anova(model.mtt.0, model.mtt.1)
```

Now we look at the CBF parameter. Considering the distribution plots, a Gamma distribution seems to be
a much better fit than a normal distribution. The $\chi_2$ test again clearly indicates that the factor "state" is significant.

```{r}
model.cbf.0 = glmer(CBF ~ roi + (1|animal), data=perfdata, family = Gamma(link = "log"))
model.cbf.1 = glmer(CBF ~ roi*state + (1|animal), data=perfdata, family = Gamma(link = "log"))
anova(model.cbf.0, model.cbf.1)
```

# Conclusion

We provide a simple package to compute perfusion parameters obtained from measurements which can be
taken using standard imaging software.

